%! Author = Len Washington III
%! Date = 10/23/24

% Preamble
\documentclass[
    % author={}, % Remove the percentage sign and add your name within the brackets.
    number={7},
]{math486homework}

% Document
\begin{document}

\maketitle

\begin{problems}
    \problem Consider the irreversible chemical reaction: $A \stackrel{k_{1}}{\rightarrow} B \stackrel{k_{2}}{\rightarrow} C$.
    \begin{problems}
       \subproblem Derive the system of differential equations for the concentrations $A = A(t)$, $B = B(t)$, and $C = C(t)$ with initial conditions $A(0) = A_{0}$, $B(0) = 0$, and $C(0) = 0$. \addanswer{Problem-1a}
       \subproblem Find a conservation law that replaces the need for the $dC/dt$ equation. \addanswer{Problem-1b}
       \subproblem Solve the equations in~\ref{prb:1a} for $A$ and $B$ then use the conservation law to find $C$. \addanswer{Problem-1c}
       \subproblem If $k_{1} = k_{2} = k$, solve for $A$, $B$, and $C$. \addanswer{Problem-1d}
       \subproblem If $k_{1} = k_{2} = k$, find the maximum amount of $B$ that is produced. \addanswer{Problem-1e}
    \end{problems}
    \problem Consider the dimerization to two monomers example from class
    \[ A + A \stackrel[k_{-1}]{k_{1}}{\rightleftarrows} C \]
    \begin{problems}
       \subproblem Derive the system of differential equations for the concentrations $A = A(t)$ and $C = C(t)$ with initial conditions $A(0) = A_{0}$ and $C(0) = 0$. \addanswer{Problem-2a}
       \subproblem Find a conservation law that replaces the need for the $dC/dt$ equation and derive the equation for $A(t)$. \addanswer{Problem-2b}
       \subproblem If $k_{1} = k_{-1} = k$ and $A_{0} = 1$, find the steady-state values for $A(t)$ and $C(t)$, i.e. limits as $t \rightarrow \infty$. \addanswer{Problem-2c}
    \end{problems}
    \subproblem Consider the ODE:
    \[ \frac{dx}{dt} = f(x) = \frac{2x^{2}}{1 + x^{4}} - x,\ \ x(0) \geq 0. \]
    \begin{problems}
       \subproblem Locate all the critical points $(x \geq 0)$. \addanswer{Problem-3a}
       \subproblem Classify the stability of all the critical points. \addanswer{Problem-3b}
       \subproblem Draw the phase line labeling the critical points and indicate the direction of flow of the solution. \addanswer{Problem-3c}
    \end{problems}
    \problem Consider the system
    \begin{equation*}
    \begin{aligned}
       \frac{dx}{dt} &= 1 - y,\\
       \frac{dy}{dt} &= x^{2} = y^{2}.
    \end{aligned}
    \end{equation*}
    \begin{problems}
       \subproblem Determine all critical points and nullclines. \addanswer{Problem-4a}
       \subproblem Find the corresponding linear system near each critical point. \addanswer{Problem-4b}
       \subproblem Find the eigenvalues and eigenvectors of each linear system.
       What conclusions can you draw about the nonlinear system? \addanswer{Problem-4c}
       \subproblem Generate a phase plane portrait for the nonlinear system.
       Use Mathematica command \textit{StreamPlot} with $-2 < x < 2$ and $-1 < y < 2$.
       Is your result in~\ref{prb:4c} consistent with the figure? \addanswer{Problem-4d}
    \end{problems}
    \problem Consider the Michaelis-Menton reaction:
    \[ S + E \stackrel[k_{-1}]{k_{1}}{\rightleftarrows} C \stackrel{k_{2}}{\rightarrow} P + E \]
    where all the rate constants are positive and the concentrations $S = $ substrate, $E = $ enzyme, $C = $ complex, and $P = $ product.
    \begin{problems}
       \subproblem Using the system of equations for $S$ and $C$ derived in class, (i.e equations (3.51) and (3.52) on page 101 of Holmes' book), find the equations of the nullclines and locate all the critical points. \addanswer{Problem-5a}
       \subproblem Test the stability of the critical points.
       Find the linearized system, compute the solution, and explain your conclusion. \addanswer{Problem-5b}
    \end{problems}
\end{problems}

\end{document}

